### 2016-04-15 ms-common-defs.tex

\definecolor{grey}{rgb}{0.5, 0.5, 0.5}
\newcommand{\smurl}[1]{\url{#1}}
\newcommand{\ud}{\ensuremath{\rm{d}}}
\newcommand{\tabitem}{~~\llap{\textbullet}~~}
\newcommand{\email}[1]{\href{mailto:#1}{\texttt{#1}}}

\title{On the link between functional traits and growth rate: meta-analysis shows effects change with plant size, as predicted}
\author{Ana{\"i}s Gibert\textasteriskcentered, Emma F. Gray, Mark Westoby, Ian J. Wright, Daniel S. Falster}
\affiliation{Department of Biological Sciences, Macquarie University,
  Sydney, Australia \\
\textasteriskcentered Correspondence author. E-mail: \texttt{anais.gibert@gmail.com}
}
\runninghead{I\lowercase{nfluence of traits on growth changes with size}}